
---
title: "Assignment 1"
output:
  html_document:
    df_print: paged
---

# A word about R markdown

The assignments for our online R training course will all take the form of R
markdown documents like this one. These are files ending in '.Rmd' that combine
R code blocks with plain text, such as this paragraph that you're reading now.

To run the R code blocks you need to have this R markdown file open in RStudio.
If you're reading this in RStudio right now then all is good and well. If not,
please start RStudio and open this file using the 'Open File...' option from the
'File' menu.

Also, it might be a good idea to create a folder on your computer in which to
store this assignment and the assignments for the coming weeks, along with some
data files that will be provided for use in some of the future assignments.

In these assignments we will describe what we're asking you to do in some plain
text which will be followed by an empty code block in which you will write the
R commands as your answers to the exercise.

Here's an example. Suppose we ask you to add two numbers, 34 and 102, together.
The empty code block where you would carry out this calculation would look like
this:

```{r}

```

It should appear as a grey box with three small grey and green icons at the top
on the right hand side.

You can fill in the answer as follows:

```{r}
34 + 102
```

You can add extra lines as necessary by hitting the 'Return' key.

Now to run this command you can click on the green arrow or triangle icon, the
one that is furthest right and points right-ward. If you hover over this icon,
you should get a tool tip saying 'Run current chunk'. Click on this and see what
happens.

You should see the answer printed just below the code block, much as you would
have seen in the Console prefixed with the (hopefully) now familiar '[1]' that
indicates this is a vector and you're looking at the first element.

In this first assignment, we are deliberately going to restrict our use of the
markdown format to just a few features. You'll notice the header containing the
title, author and date and also section headings starting with '#'. Other than
that we will just use sections of plain text and R code blocks or chunks as
they're more properly known.

As we go through successive assignments during the course, we'll introduce more
features of R markdown. It's a great way of writing R code for analyzing and
visualizing your data that lets you present your work in beautiful,
self-describing reports, a sure-fire way to impress your group leader and
colleagues alike.

R markdown is really easy to learn and will let you do a lot of cool things.
Our course website was created with R markdown documents that are not that
much more complicated than the file you're reading right now for this
assignment.

Now on with this week's exercises.

# Using R as a calculator

1. Convert the following temperatures given in degrees Fahrenheit to Celsius:
45, 96, 451

```{r}
(45-32) * 5/9
```
```{r}
(96-32) * 5/9
```


```{r}
(451-32) * 5/9
```

Hint: just do what you'd normally do if you can't remember the formula for
converting between Celsius and Fahrenheit (Google in my case).

If you like, you can experiment with getting your R code right in the Console
window first and then copy it into the code chunk above when you're happy with
it. It's not crucial and getting it wrong in the R markdown is no big deal.
You can always fix any problems (the most likely being forgetting to use
parentheses or brackets in the right place) and re-run your code using the
green arrow/triangle icon.

Check you've got the right answer by finding a web page with a handy
conversion tool.

2. Similarly, convert the following temperatures in degrees Celsius to
Fahrenheit: -65, 100, 20


```{r}
(-65 * 9/5) +32

```
```{r}
(100 * 9/5) + 32
```


```{r}
(20 * 9/5) +32
```

# Generating sequence vectors

3. Generate a sequence of numbers representing the days at which you take a
measurement or a sample at 5-day intervals for about a year.

```{r}
1:73 * 5
  

```

Your friendly neighbourhood statistician has suggested that there should be an R
function to do that. What is the function and how do you find out about it and
what is the code you will use to create the sequence? Check the resulting vector.

Answer: use Help to find information on R functions which generate a sequence 
??sequence
Answer: function below
```{r}
seq(0, 367, by = 5)
```


```{r}
```


```{r}
##notes
st <- as.Date("1998-12-17")
en <- as.Date("2000-1-7")
ll <- seq(en, st, by = "-1 month")
rev(ll[ll > st & ll < en])
```



```{r}
## notes
seq(0, 1, length.out = 11)
seq(stats::rnorm(20)) # effectively 'along'
seq(1, 9, by = 2)     # matches 'end'
seq(1, 9, by = pi)    # stays below 'end'
seq(1, 6, by = 3)
seq(1.575, 5.125, by = 0.05)
seq(17) # same as 1:17, or even better seq_len(17)

```

# Looking at types of objects

4. Run the code below and use the typeof() and/or class() function (check it's
help page) and see how R treats each newly-created vector?

```{r}
num_char <- c(1, 2, 3, "a")
num_logical <- c(1, 2, 3, TRUE)
char_logical <- c("a", "b", "c", TRUE)
tricky <- c(1, 2, 3, "4")
```

```{r}
num_char
num_logical
char_logical
tricky
```

Create a new code chunk to test each of the vectors in a separate block. You
can do this by using the 'Insert' menu just at the top of the pane for this
markdown file and selecting R for an R code chunk, or by using the keyboard
shortcut (on a Mac this is cmd-alt-i).
ANSWER: try using is. to verify an object's type  ex is.integer(days)

```{r}
??typeof
??class
```


```{r}
num_char
typeof(num_char)
class(num_char)
```
```{r}
typeof(num_char)
is.character(num_char)
```
```{r}
num_logical
typeof(num_logical)
class(num_logical)
```
```{r}
typeof(num_logical)
is.double(num_logical)
```

```{r}
char_logical
typeof(char_logical)
class(char_logical)
```


```{r}
typeof(char_logical)
is.character(char_logical)
```


```{r}
tricky
typeof(tricky)
class(tricky)
```


```{r}
typeof(tricky)
is.character(tricky)
```

You should find that R coerces the data to a lowest common denominator - can you
work out the hierarchy?

ANSWER: typeof() tells of the how the object is stored, lowest to highest common denominator is; integer, double and character.
This relates to the amount of space for storing the value and speed at accessing the value

# Plotting data

5. You have been asked to plot a graph of counts data measured over several
days. The Principal Investigator has requested that certain symbols be used for
each dataset being plotted (he's a bit like that!). What command would you use
to find out the parameter which sets this for the plot.default command and what
is the parameter's name?
Answer: args()  ex. args(plot.default) to find arguments for the function
Answer:par can be used to set or query graphical parameter as a named list 
Answer: The R function ggpubr::show_point_shapes() can be used to show the 25 commonly used R pch values.
Answer: use parameter Plot Character or pch = 0,,,25 to set or change graphing symbols 
```{r}
##install.packages("ggpubr")
ggpubr::show_point_shapes()
```

```{r}
days <- c(1, 2, 4, 6, 8, 12, 16)
counts <- days ^ 2 + rnorm(days, mean = days)

# add your code here
#ANSWER:
```


```{r}
days
```
```{r}
counts
```

```{r}
plot(counts, days, pch = 12, col = 5)
```
```{r}
??round
```

Check out what we did in the above example for getting some example counts
data points. Can you make sense of what is going on here? Look at the help
page for the `rnorm` function.

ANSWER: rnorm(x, mean =0, sd = 1) density, distribution function, quantile function and random generation for the normal distribution with mean equal to mean and standard deviation equal to sd
```{r}
days
```

Our counts data don't really look like counts as they are not whole numbers.
Find the function in R that can round these up or down to the closest whole
number and apply it in the above code chunk.
ANSWER: use help command ??round. Floor() rounds down/Ceiling() rounds up to nearest higher number
```{r}
ceiling(days)
```


```{r}
counts
floor(counts)
ceiling(counts)
```
```{r}
#using ceiling values
plot(counts, days, pch = 2, col = "red")
```

# Exploring and summarizing data

6. Your colleague has supplied you with the following table of data (number of
cells per sample volume):

| Day | LineA | LineB | LineC |
| --- | ----- | ----- | ----- |
|  1  |    4  |    5  |   14  |
|  2  |    9  |   17  |   16  |
|  3  |    7  |   22  |   10  |
|  4  |   12  |   20  |   14  |
|  5  |   23  |   24  |   20  |
|  6  |    8  |   18  |   12  |

Create some R vectors to hold this data and provide summary statistics for number
of cells for each cell line. Plot some base R graphs if you like. Describe the
data.
```{r}
Day <- c(1, 2, 3, 4, 5, 6)
Day
summary(Day)
typeof(Day)
```

```{r}
LineA <- c(4, 9, 7, 12, 23, 8)
LineA
summary(LineA)
typeof(LineA)
```
```{r}
LineB <- c(5, 17, 22, 20, 24, 18)
LineB
summary(LineB)
typeof(LineB)
```
```{r}
LineC <- c(14, 16, 10, 14, 20, 12)
LineC
summary(LineC)
typeof(LineC)
```
Plot base R graphs
```{r}
plot(LineA, Day, pch = 6, col = "black")
```
```{r}
#failed experiment at scatterplot and histogram
#scatterplot(LineA ~ Days, data=LineA)
#scatterplot(LineA ~ Days, data=LineA, smooth=list(smoother=gamLine))
#scatterplot(LineA ~ Days, data=LineA,
#    smooth=list(smoother=quantregLine))
```
```{r}
#hist(students$age)
#hist(LineA, Day)
```

You are then provided with assay data that states that LineA had an activity of
4.2 per cell, LineB an activity of 3.4 and LineC of 1.3.

Use R to calculate the activities of each sample on each day and provide summary
statistics of activity for each line.

```{r}
new_LineA <- LineA * 4.2
LineA
new_LineA

summary(new_LineA)
```
```{r}
layout(1:2)
plot(LineA, Day, pch = 7, col = "red")
plot(new_LineA, Day, pch = 7, col = "purple")
```

```{r}
new_LineB <- LineB * 3.4
LineB
new_LineB

summary(new_LineB)

```
```{r}
layout(1:2)
plot(LineB, Day, pch = 7, col = "red")
plot(new_LineB, Day, pch = 7, col = "blue")
```


```{r}
new_LineC <- LineC * 1.3
LineC
new_LineC

summary(new_LineC)
```
```{r}
layout(1:2)
plot(LineC, Day, pch = 7, col = "blue")
plot(new_LineC, Day, pch = 7, col = "red")
```


# Creating a report for your assignment

Click on the 'Knit' menu at the top of this file and select either whichever
option you prefer to create an HTML, PDF or Word document version of your
assignment. This will run all the code chunks and "knit" the resulting
results with the surrounding text to produce a report.